Crowdsourcing and phylogenetic modelling reveal parrot tool use is not rare

### 2023-08-18 manuscript.Rmd
update manuscript for peer review

keywords          : "tool use; technical intelligence; parrots; crowdsourcing; phylogenetic modelling"
wordcount         : "4273 words excluding Methods (Methods: 1792 words)"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
  - \nolinenumbers
  - \note{\raggedright † indicates shared first authorship, ‡ indicates shared senior authorship \par * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
appendix          : "appendix.Rmd"
csl               : nature.csl
---

@@

(ref:plotPhylo1Caption) _Results of crowdsourcing video survey and phylogenetic survival cure modelling mapped onto a maximum clade credibility phylogeny of the parrot order._ Purple tips indicate species observed in the scientific literature and in the video survey. Blue tips indicate species observed only in the scientific literature. Orange tips indicate species observed only in the video survey (note that three species observed only in the video survey are not present in the phylogeny due to a lack of genomic data: _Psittacara erythrogenys_, _Psittacus timneh_, and _Aratinga nenday_). Yellow tips indicate the top ten most likely tool-using species from our phylogenetic survival cure model which were not observed in the scientific literature or the video survey. All other tips are labelled as absent (grey).

@@

The identification of new tool-using species in our video survey increases the extent to which phylogeny can explain the distribution of tool use in the parrot order. We estimated phylogenetic signal (Pagel's $\lambda$) of tool use using both the pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies between 0 and 1, where 0 implies that the distribution of a trait across species is unexplained by phylogenetic relatedness and 1 implies that the distribution of a trait across species is fully explained by phylogeny. Using the evidence of tool use from the scientific literature alone (pre-video-survey data; 11 tool-using species), we estimated an average posterior Pagel's $\lambda$ of `r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` `r printnum(quantile(tar_read(phySignal1), 0.975))`]). This estimate was moderate-to-strong, but highly uncertain. In comparison, combining the evidence from both the literature and the video survey (post-video-survey data; 28 tool-using species) resulted in a stronger and more certain estimate of phylogenetic signal. With these data, we estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI [`r printnum(quantile(tar_read(phySignal2), 0.025))` `r printnum(quantile(tar_read(phySignal2), 0.975))`]). Thus, the results of our video survey increase the extent to which the distribution of tool use across parrot species can be explained by shared phylogenetic ancestry. This suggests that we can potentially use phylogenetic information, along with other variables, to identify further tool-using parrot species that may remain undetected.

@@

(ref:plotSurvCure1Caption) _Posterior predicted probabilities of tool use for each species from our phylogenetic survival cure model._ Points are posterior medians and lines are 50% and 95% credible intervals.

@@

The predicted probabilities from our phylogenetic survival cure model have implications for inferences about the evolutionary drivers and origins of tool use in the parrot order. Regarding the drivers of tool use hypothesised in Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that encephalisation quotient strongly positively predicted the probability of tool use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% CI [`r printnum(quantile(-postFull$bEQ, 0.025))` `r printnum(quantile(-postFull$bEQ, 0.975))`]; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative brain size in the dataset. We also found that feeding generalist species were slightly more likely to be tool users, though the posterior difference between generalists and specialists was quite uncertain (median posterior log odds difference = `r printnum(median(diff))`, 95% CI [`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]). These results from the survival cure model differed from the results of models fitted to pre-video-survey and post-video-survey data without the survival cure component, which found no effect of relative brain size and no difference between feeding strategies, respectively (Supplementary Figure \@ref(fig:plotComparison)).

(ref:plotSurvCure4Caption) _Posterior predictions for the effects of feeding strategy and encephalisation quotient on the probability of tool use from the phylogenetic survival cure model._ In the left plot, points and lines represent posterior medians and 50% and 95% credible intervals, with densities representing full posterior distributions. In the right plot, the line and shaded area represents the posterior median regression line with 95% credible intervals. In both plots, individual species are coloured according to the presence / absence of tool use in the video survey and the scientific literature.

@@

(ref:tableASRCaption) Estimated probabilities of tool use for most recent common ancestors of several parrot genera. _Probabilities estimated using exploratory ancestral state reconstruction models fitted to the pre-video-survey data, post-video-survey data, and predicted probabilities from the phylogenetic survival cure model._

@@

The priors for this model produce reasonable prior predictions of the probabilities of tool use for each parrot species (Supplementary Figure \@ref(fig:plotSurvCure7)), but a sensitivity analysis revealed that the ranking and posterior probabilities reported in the main text were robust to modifying these priors (Supplementary Figure \@ref(fig:plotSurvCure8)). We estimated the posterior distribution of this model using Hamiltonian Monte Carlo as implemented in Stan version 2.26.1[@Stan2020]. We iterated the model over 100 posterior parrot phylogenies to incorporate phylogenetic uncertainty. R-hat values and effective sample sizes suggested that the model converged normally. Trace plots are reported in Supplementary Figure \@ref(fig:plotTrace).

>>>>

keywords          : "tool use; technical intelligence; parrots; crowdsourcing; phylogenetic modelling"
wordcount         : "4289 words excluding Methods (Methods: 1821 words)"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
  - \nolinenumbers
  - \note{\raggedright † indicates shared first authorship, ‡ indicates shared senior authorship \par * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
appendix          : "appendix.Rmd"
csl               : nature.csl
---

@@

(ref:plotPhylo1Caption) _Results of crowdsourcing video survey and phylogenetic survival cure modelling mapped onto a maximum clade credibility phylogeny of the parrot order._ Purple tips indicate species observed in the scientific literature and in the video survey. Blue tips indicate species observed only in the scientific literature. Orange tips indicate species observed only in the video survey (note that three species observed only in the video survey are not present in the phylogeny due to a lack of genomic data: _Psittacara erythrogenys_, _Psittacus timneh_, and _Aratinga nenday_). Yellow tips indicate the top ten most likely tool-using species from our phylogenetic survival cure model which were not observed in the scientific literature or the video survey. All other tips are labelled as absent (grey). Total _n_ = `r nrow(tar_read(d))` species.

@@

The identification of new tool-using species in our video survey increases the extent to which phylogeny can explain the distribution of tool use in the parrot order. We estimated phylogenetic signal (Pagel's $\lambda$) of tool use using both the pre-video-survey and post-video-survey data. Pagel's $\lambda$ varies between 0 and 1, where 0 implies that the distribution of a trait across species is unexplained by phylogenetic relatedness and 1 implies that the distribution of a trait across species is fully explained by phylogeny. Using the evidence of tool use from the scientific literature alone (pre-video-survey data; 11 tool-using species), we estimated an average posterior Pagel's $\lambda$ of `r printnum(mean(tar_read(phySignal1)))` (95% credible interval [`r printnum(quantile(tar_read(phySignal1), 0.025))` `r printnum(quantile(tar_read(phySignal1), 0.975))`]; total _n_ = `r nrow(tar_read(d))` species). This estimate was moderate-to-strong, but highly uncertain. In comparison, combining the evidence from both the literature and the video survey (post-video-survey data; 28 tool-using species) resulted in a stronger and more certain estimate of phylogenetic signal. With these data, we estimated Pagel's $\lambda$ = `r printnum(mean(tar_read(phySignal2)))` (95% CI [`r printnum(quantile(tar_read(phySignal2), 0.025))` `r printnum(quantile(tar_read(phySignal2), 0.975))`]; total _n_ = `r nrow(tar_read(d))` species). Thus, the results of our video survey increase the extent to which the distribution of tool use across parrot species can be explained by shared phylogenetic ancestry. This suggests that we can potentially use phylogenetic information, along with other variables, to identify further tool-using parrot species that may remain undetected.

@@

(ref:plotSurvCure1Caption) _Posterior predicted probabilities of tool use for each species from our phylogenetic survival cure model._ Points are posterior medians and lines are 50% and 95% credible intervals. Total _n_ = `r nrow(tar_read(d))` species.

@@

The predicted probabilities from our phylogenetic survival cure model have implications for inferences about the evolutionary drivers and origins of tool use in the parrot order. Regarding the drivers of tool use hypothesised in Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that encephalisation quotient strongly positively predicted the probability of tool use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% CI [`r printnum(quantile(-postFull$bEQ, 0.025))` `r printnum(quantile(-postFull$bEQ, 0.975))`]; total _n_ = `r nrow(tar_read(d))` species; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative brain size in the dataset. We also found that feeding generalist species were slightly more likely to be tool users, though the posterior difference between generalists and specialists was quite uncertain (median posterior log odds difference = `r printnum(median(diff))`, 95% CI [`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]; total _n_ = `r nrow(tar_read(d))` species). These results from the survival cure model differed from the results of models fitted to pre-video-survey and post-video-survey data without the survival cure component, which found no effect of relative brain size and no difference between feeding strategies, respectively (Supplementary Figure \@ref(fig:plotComparison)).

(ref:plotSurvCure4Caption) _Posterior predictions for the effects of feeding strategy and encephalisation quotient on the probability of tool use from the phylogenetic survival cure model._ In the left plot, points and lines represent posterior medians and 50% and 95% credible intervals, with densities representing full posterior distributions. In the right plot, the line and shaded area represents the posterior median regression line with 95% credible intervals. In both plots, individual species are coloured according to the presence / absence of tool use in the video survey and the scientific literature. Total _n_ = `r nrow(tar_read(d))` species, generalist _n_ = `r sum(tar_read(d)$Feeding == "Generalist")` species, specialist _n_ = `r sum(tar_read(d)$Feeding == "Specialist")` species.

@@

(ref:tableASRCaption) Estimated probabilities of tool use for most recent common ancestors of several parrot genera. _Probabilities estimated using exploratory ancestral state reconstruction models fitted to the pre-video-survey data, post-video-survey data, and predicted probabilities from the phylogenetic survival cure model. Total_ n _= `r nrow(tar_read(d))` species._

@@

The priors for this model produce reasonable prior predictions of the probabilities of tool use for each parrot species (Supplementary Figure \@ref(fig:plotSurvCure7)), but a sensitivity analysis revealed that the ranking and posterior probabilities reported in the main text were robust to modifying these priors (Supplementary Figure \@ref(fig:plotSurvCure8)). We estimated the posterior distribution of this model using Hamiltonian Monte Carlo as implemented in Stan version 2.26.1[@Stan2020]. We ran the model for 4000 samples, with 2000 warmup samples, and iterated the model over 100 posterior parrot phylogenies to incorporate phylogenetic uncertainty. R-hat values and effective sample sizes suggested that the model converged normally. Trace plots are reported in Supplementary Figure \@ref(fig:plotTrace). We report equal-tailed credible intervals to describe the posterior distribution of this model in the main text.

### 2023-08-18 manuscript.tex
update manuscript for peer review

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{tool use; technical intelligence; parrots; crowdsourcing; phylogenetic modelling\newline\indent Word count: 4273 words excluding Methods (Methods: 1792 words)}
\usepackage{lineno}

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotPhylo1-1.pdf}
\caption{\label{fig:plotPhylo1}\emph{Results of crowdsourcing video survey and phylogenetic survival cure modelling mapped onto a maximum clade credibility phylogeny of the parrot order.} Purple tips indicate species observed in the scientific literature and in the video survey. Blue tips indicate species observed only in the scientific literature. Orange tips indicate species observed only in the video survey (note that three species observed only in the video survey are not present in the phylogeny due to a lack of genomic data: \emph{Psittacara erythrogenys}, \emph{Psittacus timneh}, and \emph{Aratinga nenday}). Yellow tips indicate the top ten most likely tool-using species from our phylogenetic survival cure model which were not observed in the scientific literature or the video survey. All other tips are labelled as absent (grey).}
\end{figure}

The identification of new tool-using species in our video survey increases the extent to which phylogeny can explain the distribution of tool use in the parrot order. We estimated phylogenetic signal (Pagel's \(\lambda\)) of tool use using both the pre-video-survey and post-video-survey data. Pagel's \(\lambda\) varies between 0 and 1, where 0 implies that the distribution of a trait across species is unexplained by phylogenetic relatedness and 1 implies that the distribution of a trait across species is fully explained by phylogeny. Using the evidence of tool use from the scientific literature alone (pre-video-survey data; 11 tool-using species), we estimated an average posterior Pagel's \(\lambda\) of 0.60 (95\% credible interval {[}0.00 0.90{]}). This estimate was moderate-to-strong, but highly uncertain. In comparison, combining the evidence from both the literature and the video survey (post-video-survey data; 28 tool-using species) resulted in a stronger and more certain estimate of phylogenetic signal. With these data, we estimated Pagel's \(\lambda\) = 0.65 (95\% CI {[}0.50 0.77{]}). Thus, the results of our video survey increase the extent to which the distribution of tool use across parrot species can be explained by shared phylogenetic ancestry. This suggests that we can potentially use phylogenetic information, along with other variables, to identify further tool-using parrot species that may remain undetected.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSurvCure1-1.pdf}
\caption{\label{fig:plotSurvCure1}\emph{Posterior predicted probabilities of tool use for each species from our phylogenetic survival cure model.} Points are posterior medians and lines are 50\% and 95\% credible intervals.}
\end{figure}

@@

The predicted probabilities from our phylogenetic survival cure model have implications for inferences about the evolutionary drivers and origins of tool use in the parrot order. Regarding the drivers of tool use hypothesised in Figure \ref{fig:plotDAG}, the phylogenetic survival cure model revealed that encephalisation quotient strongly positively predicted the probability of tool use (median posterior log odds slope = 1.12, 95\% CI {[}0.39 2.00{]}; Figure \ref{fig:plotSurvCure4}). This helps explain the ranking in Figure \ref{fig:plotSurvCure1}: the blue-eyed cockatoo has the largest relative brain size in the dataset. We also found that feeding generalist species were slightly more likely to be tool users, though the posterior difference between generalists and specialists was quite uncertain (median posterior log odds difference = 0.33, 95\% CI {[}-1.13 1.76{]}). These results from the survival cure model differed from the results of models fitted to pre-video-survey and post-video-survey data without the survival cure component, which found no effect of relative brain size and no difference between feeding strategies, respectively (Supplementary Figure \ref{fig:plotComparison}).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSurvCure4-1.pdf}
\caption{\label{fig:plotSurvCure4}\emph{Posterior predictions for the effects of feeding strategy and encephalisation quotient on the probability of tool use from the phylogenetic survival cure model.} In the left plot, points and lines represent posterior medians and 50\% and 95\% credible intervals, with densities representing full posterior distributions. In the right plot, the line and shaded area represents the posterior median regression line with 95\% credible intervals. In both plots, individual species are coloured according to the presence / absence of tool use in the video survey and the scientific literature.}
\end{figure}

@@

\caption{\label{tab:tableASR}Estimated probabilities of tool use for most recent common ancestors of several parrot genera. \emph{Probabilities estimated using exploratory ancestral state reconstruction models fitted to the pre-video-survey data, post-video-survey data, and predicted probabilities from the phylogenetic survival cure model.}}

@@

The priors for this model produce reasonable prior predictions of the probabilities of tool use for each parrot species (Supplementary Figure \ref{fig:plotSurvCure7}), but a sensitivity analysis revealed that the ranking and posterior probabilities reported in the main text were robust to modifying these priors (Supplementary Figure \ref{fig:plotSurvCure8}). We estimated the posterior distribution of this model using Hamiltonian Monte Carlo as implemented in Stan version 2.26.1\textsuperscript{80}. We iterated the model over 100 posterior parrot phylogenies to incorporate phylogenetic uncertainty. R-hat values and effective sample sizes suggested that the model converged normally. Trace plots are reported in Supplementary Figure \ref{fig:plotTrace}.

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{tool use; technical intelligence; parrots; crowdsourcing; phylogenetic modelling\newline\indent Word count: 4289 words excluding Methods (Methods: 1821 words)}
\usepackage{lineno}

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotPhylo1-1.pdf}
\caption{\label{fig:plotPhylo1}\emph{Results of crowdsourcing video survey and phylogenetic survival cure modelling mapped onto a maximum clade credibility phylogeny of the parrot order.} Purple tips indicate species observed in the scientific literature and in the video survey. Blue tips indicate species observed only in the scientific literature. Orange tips indicate species observed only in the video survey (note that three species observed only in the video survey are not present in the phylogeny due to a lack of genomic data: \emph{Psittacara erythrogenys}, \emph{Psittacus timneh}, and \emph{Aratinga nenday}). Yellow tips indicate the top ten most likely tool-using species from our phylogenetic survival cure model which were not observed in the scientific literature or the video survey. All other tips are labelled as absent (grey). Total \emph{n} = 174 species.}
\end{figure}

The identification of new tool-using species in our video survey increases the extent to which phylogeny can explain the distribution of tool use in the parrot order. We estimated phylogenetic signal (Pagel's \(\lambda\)) of tool use using both the pre-video-survey and post-video-survey data. Pagel's \(\lambda\) varies between 0 and 1, where 0 implies that the distribution of a trait across species is unexplained by phylogenetic relatedness and 1 implies that the distribution of a trait across species is fully explained by phylogeny. Using the evidence of tool use from the scientific literature alone (pre-video-survey data; 11 tool-using species), we estimated an average posterior Pagel's \(\lambda\) of 0.60 (95\% credible interval {[}0.00 0.90{]}; total \emph{n} = 174 species). This estimate was moderate-to-strong, but highly uncertain. In comparison, combining the evidence from both the literature and the video survey (post-video-survey data; 28 tool-using species) resulted in a stronger and more certain estimate of phylogenetic signal. With these data, we estimated Pagel's \(\lambda\) = 0.65 (95\% CI {[}0.50 0.77{]}; total \emph{n} = 174 species). Thus, the results of our video survey increase the extent to which the distribution of tool use across parrot species can be explained by shared phylogenetic ancestry. This suggests that we can potentially use phylogenetic information, along with other variables, to identify further tool-using parrot species that may remain undetected.

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSurvCure1-1.pdf}
\caption{\label{fig:plotSurvCure1}\emph{Posterior predicted probabilities of tool use for each species from our phylogenetic survival cure model.} Points are posterior medians and lines are 50\% and 95\% credible intervals. Total \emph{n} = 174 species.}
\end{figure}

@@

The predicted probabilities from our phylogenetic survival cure model have implications for inferences about the evolutionary drivers and origins of tool use in the parrot order. Regarding the drivers of tool use hypothesised in Figure \ref{fig:plotDAG}, the phylogenetic survival cure model revealed that encephalisation quotient strongly positively predicted the probability of tool use (median posterior log odds slope = 1.12, 95\% CI {[}0.39 2.00{]}; total \emph{n} = 174 species; Figure \ref{fig:plotSurvCure4}). This helps explain the ranking in Figure \ref{fig:plotSurvCure1}: the blue-eyed cockatoo has the largest relative brain size in the dataset. We also found that feeding generalist species were slightly more likely to be tool users, though the posterior difference between generalists and specialists was quite uncertain (median posterior log odds difference = 0.33, 95\% CI {[}-1.13 1.76{]}; total \emph{n} = 174 species). These results from the survival cure model differed from the results of models fitted to pre-video-survey and post-video-survey data without the survival cure component, which found no effect of relative brain size and no difference between feeding strategies, respectively (Supplementary Figure \ref{fig:plotComparison}).

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/plotSurvCure4-1.pdf}
\caption{\label{fig:plotSurvCure4}\emph{Posterior predictions for the effects of feeding strategy and encephalisation quotient on the probability of tool use from the phylogenetic survival cure model.} In the left plot, points and lines represent posterior medians and 50\% and 95\% credible intervals, with densities representing full posterior distributions. In the right plot, the line and shaded area represents the posterior median regression line with 95\% credible intervals. In both plots, individual species are coloured according to the presence / absence of tool use in the video survey and the scientific literature. Total \emph{n} = 174 species, generalist \emph{n} = 121 species, specialist \emph{n} = 53 species.}
\end{figure}

@@

\caption{\label{tab:tableASR}Estimated probabilities of tool use for most recent common ancestors of several parrot genera. \emph{Probabilities estimated using exploratory ancestral state reconstruction models fitted to the pre-video-survey data, post-video-survey data, and predicted probabilities from the phylogenetic survival cure model. Total} n \emph{= 174 species.}}

@@

The priors for this model produce reasonable prior predictions of the probabilities of tool use for each parrot species (Supplementary Figure \ref{fig:plotSurvCure7}), but a sensitivity analysis revealed that the ranking and posterior probabilities reported in the main text were robust to modifying these priors (Supplementary Figure \ref{fig:plotSurvCure8}). We estimated the posterior distribution of this model using Hamiltonian Monte Carlo as implemented in Stan version 2.26.1\textsuperscript{80}. We ran the model for 4000 samples, with 2000 warmup samples, and iterated the model over 100 posterior parrot phylogenies to incorporate phylogenetic uncertainty. R-hat values and effective sample sizes suggested that the model converged normally. Trace plots are reported in Supplementary Figure \ref{fig:plotTrace}. We report equal-tailed credible intervals to describe the posterior distribution of this model in the main text.
